### 2018-06-12 implicit.tex

\subsection{Motivation}

The motivation behind implicit morphisms is to generalize the benefits of include declarations even further.
Sometimes an atomic morphism is conceptually similar to an inclusion: For example, we might want to declare $m:\cn{Division}\to\cn{Group}=\{\cn{divide} := \lambda x,y.x\circ y^{-1}, \;\ldots\}$ to show how division is ``included'' into the theory of groups.
Other times we have to use structures because we need to duplicate declarations but we would still like to use them like includes.
For example, we have to declare the theory of rings using two structures (for the commutative group and the monoid).
But we still want a ring to be implicitly converted into, e.g., a commutative group.

Our key idea is to use a commutative{\footnotemark} subdiagram of the \mmt diagram, which we call the \emph{implicit-diagram}.
It contains all theories but only some of the morphisms -- the ones designated as \emph{implicit}.
Because the implicit-diagram commutes, there can be at most one implicit morphism from $S$ to $T$ -- if this morphism is $M$, we write $\ipc{M}{S}{T}$.
\footnotetext{A diagram commutes if for any two morphisms $M,M'\in\Mo{S}{T}$, we have $M\equiv M'$.}
%Practically, this hasis a multi-graph whose nodes are the theories and whose edges are the atomic morphisms.
%We can extend the correspondence by identifying all paths with the composition of all edges along the path.
%Then concatenation of paths is just composition of morphisms, and the empty path from $T$ to itself is the identity morphism $\id{T}$.
%(This is well-defined because composition is associative and identity is neutral.)
%A diagram is commutative if any two paths between the same nodes are equal (with respect to $\equiv$), i.e., there is at most one morphism between any two nodes.

The implicit-diagram generalizes the inclusion relation $\harr$.
In particular, all identity and inclusion morphisms are implicit, and we recover $S\harr T$ as the special case $\ipc{\id{S}}{S}{T}$.
Moreover, the relation ``exists $M$ such that $\ipc{M}{S}{T}$'' is also an order.

Consequently, many of the advantages of inclusions carry over to implicit morphisms.
\begin{compactitem}
\item It is very easy to maintain the implicit-diagram, e.g., as a partial map that assigns to a pair of theories the implicit morphism between them (if any).
\item We can generalize the visibility of identifiers: If $\ipc{M}{S}{T}$, we can use all $S$-identifiers in $T$ as if $S$ were included into $T$.
Any $c\in\dom{S}$ is treated as a valid $T$-identifiers with definiens $M(c)$.
\item We can still use canonical identifiers $s?n$. Because there can be at most one implicit morphism $\ipc{M}{s}{T}$, using $s?n$ as an identifier in $T$ is unambiguous.
Crucially, we can use $s?n$ without bothering what $M$ is.
\end{compactitem}

%Thus the main question arises what other morphisms we should make implicit.

\subsection{Definition}

\paragraph{Syntax}
We have already indicated that we will consider all identities and inclusions as implicit.
It remains to give users the option introduce additional implicit morphisms.
For that purpose, we allow two \mmt declarations to carry an attribute: 
\begin{grammar}
MDec   & Att\; m:T\to T=\{Ass,\ldots,Ass\}   & \text{attributed atomic morphism} \\
Dec    & Att\; \icl{S}                       & \text{attributed inclusion}\\
Att    & \keyword{implicit} \alt \keyword{invertible} & \text{attributes}
\end{grammar}

Every morphism with the attribute \keyword{implicit} occurs in the implicit-diagram.
(Of course, this is redundant for includes.)
Every morphism with the attribute \keyword{invertible} is also considered implicit, and additionally so is its inverse morphism.

We do not even have to provide a name for the inverse morphism: It is uniquely determined (up to $\equiv$) anyway, and it will be used only implicitly.
We only have to make sure that the inverse actually exists and that it can be found automatically.
This is not decidable in general, but it is decidable in many important special cases, in particular, renamings and definitional extensions.

\paragraph{Semantics}
Well-formedness and semantics are defined in parallel.

\begin{definition}[Implicit-Diagram]
Given a diagram $D$, its implicit-diagram $D^i$ contains all theories and the following morphisms:
\begin{compactitem}
\item all identities and inclusions,
\item all compositions of implicit morphisms,
\item every morphism with the \keyword{implicit} attribute,
\item for each morphism with the \keyword{invertible} attribute, that morphism and its inverse. 
\end{compactitem}
\end{definition}

The only requirement for well-formedness is that the implicit-diagram always commutes.
However, in general, commutativity may be undecidable.
Similarly, the existence of an inverse morphism may be undecidable.
Therefore, we use sound criteria for commutativity and invertibility as defined in Sect.~\ref{sec:commute} and \ref{sec:inverse}.

\begin{definition}[Well-Formedness]
Consider a well-formed diagram $D, m:S\to T=\{\sigma\}$.
Let $D^a$ be $D, a\; m:S\to T=\{\sigma\}$.
$D^{\keyword{implicit}}$ is well-formed if $D^i$ extended with an edge for $m$ commutes.
$D^{\keyword{invertible}}$ is well-formed if $D^i$ extended with edges for $m$ and its inverse commutes.

Correspondingly, consider a well-formed diagram $D$ of the form $D_0, T=\{\ldots,\icl{S},\ldots\}$.
Let $D^a$ be $D_0, T=\{\ldots, a\; \icl{S},\ldots\}$.
$D^{\keyword{invertible}}$ is well-formed if the inclusion has an *inverse.
\end{definition}

Finally, we modify the definition of $\Exp{T}$ and $M(-)$ to add all implicitly imported identifiers:

\begin{definition}[Visible Declarations]
The set $\flti{T}$ contains the following declarations:
for every $\ipc{M}{S}{T}$ and every $c[:t][=d]$ in $\flt{S}$, the declaration $c[:M(t)]=M(c)$.

The set $\Expi{T}$ contains all objects formed from the symbols in $\flti{T}$.
The map $M(-)$ maps every $c$ for which $\flt{M}$ does not provide an assignment to $M(d)$ where $d$ is the definiens of $c$.
\end{definition}

This modification is conservative in the sense that $\flt{T}\sq\flti{T}$ and $\flti{T}\sm\flt{T}$ contains only declarations with definiens.
Moreover, $\Exp{T}\sq\Expi{T}$ and $M(-)$ remains unchanged on all $o\in\Exp{T}$.
It is easy to see that $M(-)$ still preserve all judgments.

\subsection{Commutativity}\label{sec:commute}

To prove commutativity, we have to find all paths in the implicit-diagram and check $M\equiv M'$ for all pairs of morphisms between the same nodes.

\mmt reduces $M\equiv M'$ for $M,M'\in\Mo{S}{T}$ to a first-order problem in the theory of categories, which uses the following axioms:
\begin{compactitem}
\item associativity of composition,
\item neutrality of identity,
\item $M;M'\equiv \id{S}$ as well as $M';M\equiv\id{T}$ if $M\in\Mo{S}{T}$ and $M'$ is the known *inverse of $M$,
\item if an atomic morphism $m$ includes $M\in\Mo{R}{T}$, then $M|_R\equiv M$.
\end{compactitem}

This is a sound but not complete axiomatization of $M\equiv M'$.
We can obtain a complete axiomatization in a totally different way: Check $\der_T M(c)\equiv M'(c)$ for all $c\in\dom{S}$.
However, this can be extremely expensive: It requires computing $\dom{S}$ and proving an object level equality for each identifier.
Even if the equality of objects is decidable, this is often too expensive.

In fact, because implicit morphisms are meant to be used in very particular cases, we expect the above incomplete calculus to perform reasonably well in practice.
If showing the equality of certain morphisms is hard, one should probably not make them implicit in the first place.

\subsection{Inverse Morphisms}\label{sec:inverse}

We only two use sufficient criteria to determine an *inverse of a morphism $M$.

\paragraph{Conservative Extensions}
Intuitively, $t$ extends $S$ definitionally, if it only adds defined symbols, e.g., abbreviations or theorems.
That is the most important case of conservative extensions.

\begin{definition}[Definitional Extension]
An include declaration \[t=\{\keyword{invertible}\;\icl{S},\Sigma\}\] is \textbf{definitional} if
\begin{compactitem}
 \item all symbol declarations in $\Sigma$ have a definiens,
 \item all include declarations $\icl{R}$ in $\Sigma$ are such that $R$ also definitionally includes $S$.
\end{compactitem}
\end{definition}

There are multiple, subtly different definitions of conservative extension.
However, definitional extensions as defined above are always conservative (i.e., for any reasonable definition of conservativity and in any reasonable formal language).
In fact, they are always invertible: The inverse morphism $i:t\to S$ maps
\begin{compactitem}
 \item an identifier $c\in\dom{S}$ to $c$ and everythemselves and the others to their definiens,
 \item any other identifier, which must have a definiens $d$, to $i(d)$.
\end{compactitem}

%We use two auxiliary definitions.
%Given $u:T\to U$, we write $u|_S$ for the restriction of $u$ to $S$, i.e., the composition of the inclusion $S\harr T$ and $u$.
%An extension $S\to T$ is retractable if there is a morphism $r:T\to S$ such that $r|_S=\id{S}$.
%
%Whether a given extension $S\harr T$ is conservative depends on the specific definition of conservativity and the used logic.
%However, the most important kind of conservative extension is very easy to handle: the extension of a theory with a defined constant (e.g., an abbreviation or a theorem).
%These extensions are always conservative (i.e., for any reasonable definition of conservativity and in any reasonable logic).
%In fact, they are always isomorphic to the original theory.
%
%Extensions with definition are almost exclusively used to add something to $S$.
%Therefore, it is reasonable to automatically designate the retraction as implicit.
%In particular, this provides an elegant way to add theorems to $S$ in such a way that they are available to any theory importing $S$, but without actually changing $S$.
%
%More generally, any retractable extension $S\harr T$ is always conservative.
%But if the retraction usually does not exist uniquely (Often an arbitrary non-canonical choice is needed, e.g., any type.), this is not enough to make the retraction implicit.

\paragraph{Renamings}
Intuitively, an atomic morphism $m:S\to T$ is a renaming if it maps all $S$-identifiers to $T$-identifiers (rather than $T$-objects).

\begin{definition}[Renaming]
Consider a morphism $M\in\Mo{S}{T}$.
Let $A\sq\dom{S}$ and $B\sq\dom{T}$ contain those identifiers without definiens.
$M$ is an (injective, surjective) \textbf{renaming} if the restriction of $M(-)$ is an (injective, surjective) mapping $A\to B$.
\end{definition}

Clearly, if $m$ is an injective and surjective renaming, it is straightforward to obtain its inverse by inverting $M(-)$.

\subsection{Applications}

Canonical isomorphisms.

User-declared atomic morphisms.

\paragraph{Extending Theories from the Outside}

\paragraph{Transparent Refactoring}
A major drawback of OpenMath/\mmt-style qualified identifiers is that it can preclude transparent refactoring.
For example, consider a theory $t=\{\icl{r},\icl{s}\}$ and assume we want to move a symbol $n$ from $r$ to $s$.
This requires changing any qualified reference to $r?n$ (in any theory including $t$) to $s?n$.
That can be very costly, especially in large libraries where the refactorer might not even know of or have write access to all theories importing $t$.

With implicit morphisms, we rename $s$ to $s'$, move $n$ from $s'$ to $r$, and recover $s$ by $s=\{\icl{s'},n\}$.
Now we can change define $t=\{\icl{r},\icl{s'}\}$ and give an implicit morphism from $s$ to $t$, which maps $s?n$ to $r?n$.
All existing references to $s?n$ stay well-formed (and serve as aliases for $r?n$) so that no further changes in theories importing $t$ are needed.