### 2018-06-12 intro.tex

The contribution of this paper is threefold.

Firstly, we describe the syntax and semantics of include declarations in \mmt.
Our main interest here is that inclusion are the most important form of implicit morphisms.
However, this is also the first item that the semantics of includes in MMT is specified rigorously.

Secondly, we generalize inclusions to implicit morphisms and show how we can use implicit morphisms to structure large diagrams of theories.
Most notably, we show that many of the goals of realms \cite{realms} (which won the best paper award at MKM 2014) are already realized by \mmt in the presence of implicit morphisms.

Thirdly, we give an extension of \mmt with complex theory and morphism expressions.
Our focus here is on defining which of these complex morphisms can be designated as implicit.
However, the language is novel and a valuable contribution in itself.