### 2018-06-12 paper.tex

\documentclass[orivec]{llncs}

\usepackage{graphicx,url,wrapfig,amsmath,amssymb}

\usepackage[show]{ed} %for ednotes

\usepackage{sembracket}
\usepackage{mytikz}
\usepackage{basics}
%\usepackage[llncs]{theorems} %only for now so that the environments are defined; replace with nice package for this later
\usepackage{local}

\setcounter{tocdepth}{3}
\usepackage[bookmarks,linkcolor=red,citecolor=blue,urlcolor=gray,colorlinks,breaklinks,bookmarksopen,bookmarksnumbered]{hyperref}

\setlength{\hfuzz}{3pt}
\hbadness=10001 %make box warning less strict

\begin{document}
\title{Structuring Theories with Implicit Morphisms}
\author{Florian Rabe$^{1,2}$ \and Dennis M\"uller$^2$}
\institute{LRI Paris \and FAU Erlangen-Nuremberg}
\maketitle

\begin{abstract}
We introduce \emph{implicit} morphisms as a concept in formal systems based on theories and theory morphisms.
The idea is that there may be at most one implicit morphism from a theory $S$ to a theory $T$, and if $S$-expressions are used in $T$ their semantics is obtained by automatically inserting the implicit morphism.
The practical appeal of implicit morphisms is that they hit the sweet-spot of being extremely simple to understand and implement while significantly helping with structuring large collections of theories.
\end{abstract}

\input{wadtabstract}
\bibliographystyle{alpha}
\bibliography{../../../Program_Data/Latex/bib/rabe,../../../Program_Data/Latex/bib/systems,../../../Program_Data/Latex/bib/institutions,../../../Program_Data/Latex/bib/pub_rabe,../../../Program_Data/Latex/bib/historical}
\end{document}

\section{Introduction}
\input{intro}

\section{Theories and Theory Morphism}
\input{mmt}

\section{Inclusion Morphisms}\label{sec:include}
\input{include}

\section{Implicit Morphisms}
\input{implicit}

\section{Complex Theories and Morphisms}\label{sec:complex}
\input{complex}

\section{Related Work}
\input{related}

\section{Conclusion}
\input{conc}

\bibliographystyle{alpha}
\bibliography{../../../Program_Data/Latex/bib/rabe,../../../Program_Data/Latex/bib/systems,../../../Program_Data/Latex/bib/institutions,../../../Program_Data/Latex/bib/pub_rabe,../../../Program_Data/Latex/bib/historical}

\end{document}