### 2018-09-27 example.tex
Bands example

A good example for the usefulness of implicit morphisms is the lattice of \emph{bands}. A band is an idempotent semigroup. There are several varieties of bands which can all be defined by a single governing equation, resulting in the lattice in \autoref{fig:bands}.

\begin{figure}\begin{center}
\includegraphics[width=0.6\textwidth]{bands}\end{center}
\caption{The Lattice of Varieties of Bands}\label{fig:bands}
\end{figure}

As the diagram suggests, even though each node in the lattice is definable via a single equation, each of them implies the connected equations above it. Hence formalizing the varieties of bands via includes would add the implied equations as \emph{additional axioms}, even though they are provable.

Using implict morphisms between the varieties however can turn the axiom in a variety of bands into a theorem in another variety. \autoref{fig:bandsmmt} shows a subset of this lattice implemented with implict morphisms.

\begin{figure}
\begin{tabular}{c|c}
\begin{mmtcode}
theory Band : ?Meta =
  include ?SemiGroup ❙	
  axiom_idemp : ⊦ ∀[x] x ∘ x ≐ x ❙
❚

theory Regular : ?Meta =
  include ?Band ❙
  axiom_regular : ⊦ ∀[x]∀[y]∀[z] 
    z ∘ x ∘ z ∘ y ∘ z ≐ z ∘ x ∘ y ∘ z ❙
❚

theory LeftNormal : ?Meta =
  include ?Band ❙
  axiom_leftnormal : ⊦ ∀[x]∀[y]∀[z] 
    z ∘ x ∘ z ∘ y ≐ z ∘ x ∘ y ❙ 
❚

theory RightNormal : ?Meta =
  include ?Band ❙
  axiom_rightnormal : ⊦ ∀[x]∀[y]∀[z] 
    y ∘ z ∘ x ∘ z ≐ y ∘ x ∘ z ❙ 
❚

theory Normal : ?Meta =
  include ?Band ❙
  axiom_normal :  ⊦ ∀[x]∀[y]∀[z] 
    z ∘ x ∘ y ∘ z ≐ z ∘ y ∘ x ∘ z ❙
❚
\end{mmtcode} &
\begin{mmtcode}
implicit view Reg2LeftNormal : 
    ?Regular -> ?LeftNormal =
  include ?Band = ?Band ❙
  axiom_regular = sketch "trivial" ❙
❚

implicit view Reg2RightNormal : 
    ?Regular -> ?RightNormal =
  include ?Band = ?Band ❙
  axiom_regular = sketch "trivial" ❙
❚

implicit view LeftNormal2Normal : 
    ?LeftNormal -> ?Normal =
  include ?Band = ?Band ❙
  axiom_leftnormal = sketch "trivial" ❙
❚

implicit view RightNormal2Normal : 
    ?RightNormal -> ?Normal =
  include ?Band = ?Band ❙
  axiom_rightnormal = sketch "trivial" ❙
❚
\end{mmtcode}
\end{tabular}

\caption{Exemplary Theories for Some Varieties of Bands}\label{fig:bandsmmt}
\end{figure}