### 2018-10-14 paper.tex

\documentclass[orivec]{llncs}

\usepackage{graphicx,url,wrapfig,amsmath,amssymb}

\usepackage{comment}
% flip these to add module expressions; if so, some manual changes to intro and grammars are also necessary
\excludecomment{modexp}
\includecomment{nomodexp}
% include unions as an example of module expressions
\excludecomment{union}

\usepackage[show]{ed} %for ednotes

\usepackage{macros/sembracket}
\usepackage{macros/mytikz}
\usepackage{macros/basics}
\usepackage{local}

\input{macros/listings}

\setcounter{tocdepth}{3}
\usepackage[bookmarks,linkcolor=red,citecolor=blue,urlcolor=gray,colorlinks,breaklinks,bookmarksopen,bookmarksnumbered]{hyperref}

\setlength{\hfuzz}{3pt}
\hbadness=10001 %make box warning less strict

\pagestyle{plain}

\begin{document}
\title{Structuring Theories with Implicit Morphisms}
\author{Florian Rabe$^{1,2}$ \and Dennis M\"uller$^2$}
\institute{LRI Paris \and FAU Erlangen-Nuremberg}
\maketitle

\begin{abstract}
We introduce \emph{implicit} morphisms as a concept in formal systems based on theories and theory morphisms.
The idea is that there may be at most one implicit morphism from a theory $S$ to a theory $T$, and if $S$-expressions are used in $T$ their semantics is obtained by automatically inserting the implicit morphism.
The practical appeal of implicit morphisms is that they hit the sweet-spot of being extremely simple to understand and implement while significantly helping with structuring large collections of theories.
\end{abstract}

\section{Introduction}
\input{intro}

\section{Theories and Theory Morphism}\label{sec:mmt}
\input{mmt}

%\section{Inclusion Morphisms}\label{sec:include}
%\input{include}

\section{Implicit Morphisms}\label{sec:impl}
\input{implicit}

\section{Applications}\label{sec:appl}
\input{appl}

\begin{modexp}
\section{Advanced Module Expressions}\label{sec:complex}
\input{complex}
\end{modexp}

\section{Conclusion and Related Work}\label{sec:conc}
\input{conc}

\bibliographystyle{alpha}
\bibliography{../../../../frabe/Program_Data/Latex/bib/rabe,../../../../frabe/Program_Data/Latex/bib/systems,../../../../frabe/Program_Data/Latex/bib/institutions,../../../../frabe/Program_Data/Latex/bib/pub_rabe,../../../../frabe/Program_Data/Latex/bib/historical}

\end{document}