### 2018-10-19 macros/listings.tex

\usepackage{listings}

\lstdefinelanguage{mmt}{
	morekeywords={theory,namespace,rule,role,import,implicit,definitional,renaming},
	morecomment=[l][\color{Gray}]{//},
	% morecomment={[s]{//}{❚},[s]{//}{❙},[s]{//}{❘}}
}
\lstnewenvironment{mmtcode}[1][language=mmt]{\lstset{language=mmt,keywordstyle=\bfseries,basicstyle=\ttfamily,mathescape,columns=fixed,#1}}{}