### 2019-02-11 paper.tex

\documentclass[orivec]{llncs}

\usepackage{graphicx,url,wrapfig,amsmath,amssymb}

\usepackage{comment}
% flip these to add module expressions; if so, some manual changes to intro and grammars are also necessary
\excludecomment{modexp}
\includecomment{nomodexp}
% include unions as an example of module expressions
\excludecomment{union}

\usepackage[show]{ed} %for ednotes

\usepackage{macros/sembracket}
\usepackage{macros/mytikz}
\usepackage{macros/basics}
\usepackage{local}

\input{macros/listings}

\setcounter{tocdepth}{3}
\usepackage[bookmarks,linkcolor=red,citecolor=blue,urlcolor=gray,colorlinks,breaklinks,bookmarksopen,bookmarksnumbered]{hyperref}

\setlength{\hfuzz}{3pt}
\hbadness=10001 %make box warning less strict

\pagestyle{plain}

\begin{document}
\title{Structuring Theories with Implicit Morphisms}
\author{Florian Rabe$^{1,2}$ \and Dennis M\"uller$^2$}
\institute{LRI Paris \and FAU Erlangen-Nuremberg}
\maketitle

\begin{abstract}
We introduce \emph{implicit} morphisms as a concept in formal systems based on theories and theory morphisms.
The idea is that there may be at most one implicit morphism from a theory $S$ to a theory $T$, and if $S$-expressions are used in $T$ their semantics is obtained by automatically inserting the implicit morphism.
The practical appeal of implicit morphisms is that they hit the sweet-spot of being extremely simple to understand and implement while significantly helping with structuring large collections of theories.
Concrete applications include elegantly identifying isomorphic theories and extending theories with definitions and theorems as well as efficiently building and maintaining large, fine-granular, and heterogeneous hierarchies of theories.
\end{abstract}

\section{Introduction}
\input{intro}

\section{Theories and Theory Morphisms}\label{sec:mmt}
\input{mmt}

%\section{Inclusion Morphisms}\label{sec:include}
%\input{include}

\section{Implicit Morphisms}\label{sec:impl}
\input{implicit}

\section{Applications}\label{sec:appl}
\input{appl}

\begin{modexp}
\section{Advanced Module Expressions}\label{sec:complex}
\input{complex}
\end{modexp}

\section{Conclusion and Related Work}\label{sec:conc}
\input{conc}

\bibliographystyle{alpha}
\bibliography{macros/pub_rabe,macros/rabe,macros/systems,macros/institutions}

\end{document}